Overview and Summary of the First Workshop on Sustainable Software for Science: Practice and Experiences (WSSSPE1) \note{\scriptsize Authors, please add/check middle initials and affiliations}

### 2014-04-14 WSSSPE1_report/WSSSPE1_summary_report.tex
Acknowledgement.
Fix et al.

One group of papers presented frameworks that were primarily about
characteristics of software artifacts, connecting with the long
discourse on software quality. This approach is realized in adjectives
that can be applied to pieces of software but might also extend to
describe software projects.  Thus Calero et al.~\cite{Calero_WSSSPE} propose adding elements to the ISO
standards for measuring software quality. Specifically, they propose an additional
dimension of quality they call ``perdurability'' or ability to ``last over
time'' made up of three characteristics: reliability, maintainability
and adaptability. This shows some overlap with the framework employed
by Venters et al.~\cite{Venters_WSSSPE} who employ the adjectives of
``extensibility, interoperability, maintainability, portability,
reusability and scalability,'' anticipating the sorts of work that
people would need to do with a software artifact in the future, ``as
stakeholders requirements, technology and environments evolve and
change.'' Venters et al. argue that these choices are really made very
early because they relate to the architecture of the software and
involve trade offs that ought to be analyzed alongside other
attributes of the software artifact. Lenhardt et
al~\cite{Lenhardt_WSSSPE} compare the software lifecycle to the data
lifecycle to argue for the inclusion of metadata throughout a piece of
software's life (discussing, for example, how it has been built and
tested and what ``data in'' and ``data out'' has been considered). In
addition, their analogy suggests that the software lifecycle might add
a phase of ``preservation'' and draw on understandings of what that
involves from studies of data.  In sum, then, these frameworks focused
on what needs to be accomplished to have more sustainable software.

@@

How work on scientific software is recognized and rewarded strongly influences the motivation for particular kinds of work
on scientific software. A recurring theme of the panel discussion was
that software work in science is
inadequately visible within the reputation
system underlying science; in other words it often doesn't ``count''. In his paper for this workshop, Katz placed
software work along with other ``activities that facilitate science
but are not currently rewarded or
recognized''~\cite{Katz_WSSSPE}. Priem and Piwowar argued for the need
to ``support all researchers in presenting meaningful impact evidence
in tenure, promotion, and funding applications.''~\cite{Priem_WSSSPE}.
Knepley et al. argued that the lack of visibility of software that supported a
piece of science ``can have detrimental effects on funding, future
library development, and even scientific
careers.''~\cite{Knepley_WSSSPE}.

@@

Krintz et al.~\cite{Krintz_WSSSPE} describe how in an era in which
computing frequently takes place in a distributed cloud, the control
over digital resources is increasingly shifting from physical
infrastructure to APIs, in particular web-service APIs. Yet, as Krintz
et al. observe, unlike for physical IT infrastructure in data centers,
science communities have developed very little in the way of practices and
technology for API governance, referred to by Krintz et al. as the
``combined policy, implementation, and deployment control''. Web APIs
can and do change, sometimes quite frequently. Raising the need to
port dependent applications. The effort required for porting is
notoriously difficult to estimate, making it nearly impossible for IT
organizations to assess and thus properly manage the impact of API
changes. To address this, Krintz et al. propose a mechanism that
evaluates the porting effort between two versions of a web-service API
in a formal and automated way. To analyze the porting effort, they
divide API changes into \emph{syntactic similarity}, the changes in
inputs and outputs, and into \emph{semantic similarity}, the changes
in the API's behavior. In initial tests, their method showed good
congruence with human developers in scoring porting effort, offering
the possibility that API governance can become as solid a part of
scientific IT management as data center infrastructure management is
today.

@@

>>>>

One group of papers presented frameworks that were primarily about
characteristics of software artifacts, connecting with the long
discourse on software quality. This approach is realized in adjectives
that can be applied to pieces of software but might also extend to
describe software projects.  Thus Calero et al.~\cite{Calero_WSSSPE} propose adding elements to the ISO
standards for measuring software quality. Specifically, they propose an additional
dimension of quality they call ``perdurability'' or ability to ``last over
time'' made up of three characteristics: reliability, maintainability
and adaptability. This shows some overlap with the framework employed
by Venters et al.~\cite{Venters_WSSSPE} who employ the adjectives of
``extensibility, interoperability, maintainability, portability,
reusability and scalability,'' anticipating the sorts of work that
people would need to do with a software artifact in the future, ``as
stakeholders requirements, technology and environments evolve and
change.'' Venters et al.~argue that these choices are really made very
early because they relate to the architecture of the software and
involve trade offs that ought to be analyzed alongside other
attributes of the software artifact. Lenhardt et
al.~\cite{Lenhardt_WSSSPE} compare the software lifecycle to the data
lifecycle to argue for the inclusion of metadata throughout a piece of
software's life (discussing, for example, how it has been built and
tested and what ``data in'' and ``data out'' has been considered). In
addition, their analogy suggests that the software lifecycle might add
a phase of ``preservation'' and draw on understandings of what that
involves from studies of data.  In sum, then, these frameworks focused
on what needs to be accomplished to have more sustainable software.

@@

How work on scientific software is recognized and rewarded strongly influences the motivation for particular kinds of work
on scientific software. A recurring theme of the panel discussion was
that software work in science is
inadequately visible within the reputation
system underlying science; in other words it often doesn't ``count''. In his paper for this workshop, Katz placed
software work along with other ``activities that facilitate science
but are not currently rewarded or
recognized''~\cite{Katz_WSSSPE}. Priem and Piwowar argued for the need
to ``support all researchers in presenting meaningful impact evidence
in tenure, promotion, and funding applications.''~\cite{Priem_WSSSPE}.
Knepley et al.~argued that the lack of visibility of software that supported a
piece of science ``can have detrimental effects on funding, future
library development, and even scientific
careers.''~\cite{Knepley_WSSSPE}.

@@

Krintz et al.~\cite{Krintz_WSSSPE} describe how in an era in which
computing frequently takes place in a distributed cloud, the control
over digital resources is increasingly shifting from physical
infrastructure to APIs, in particular web-service APIs. Yet, as Krintz
et al.~observe, unlike for physical IT infrastructure in data centers,
science communities have developed very little in the way of practices and
technology for API governance, referred to by Krintz et al.~as the
``combined policy, implementation, and deployment control''. Web APIs
can and do change, sometimes quite frequently. Raising the need to
port dependent applications. The effort required for porting is
notoriously difficult to estimate, making it nearly impossible for IT
organizations to assess and thus properly manage the impact of API
changes. To address this, Krintz et al.~propose a mechanism that
evaluates the porting effort between two versions of a web-service API
in a formal and automated way. To analyze the porting effort, they
divide API changes into \emph{syntactic similarity}, the changes in
inputs and outputs, and into \emph{semantic similarity}, the changes
in the API's behavior. In initial tests, their method showed good
congruence with human developers in scoring porting effort, offering
the possibility that API governance can become as solid a part of
scientific IT management as data center infrastructure management is
today.

@@

Choi thanks Ian Foster, Director of the Computation Institute, University of Chicago,
for encouraging her work in reliable reproducible research and 
supporting her trip to participate in the WSSSPE1 conference as a contributor.
